Taking Linear Logic Apart

### 2019-07-13 doc/paper/main.tex
minor fixes

\subsection{Metatheory}
\cp enjoys subject reduction, termination, and progress~\cite{wadler2012,lindley2015}.
\begin{lemma}[Preservation for $\equiv$]\label{lem:cp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\Gamma}$ iff $\seq[Q]{\Gamma}$.
\end{lemma} 
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:cp-preservation}
  If $\seq[P]{\Gamma}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\Gamma}$.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}
\begin{definition}[Actions]
  A process $\tm{P}$ acts on $\tm{x}$ whenever $\tm{x}$ is free in the outermost
  term constructor of $\tm{P}$, \eg $\tm{\cpSend{x}{y}{P}{Q}}$ acts on $\tm{x}$
  but not on $\tm{y}$, and $\tm{\cpLink{x}{y}}$ acts on both $\tm{x}$ and $\tm{y}$.
  A process $\tm{P}$ is an action if it acts on some channel $\tm{x}$.
\end{definition}
\begin{definition}[Canonical forms]
  A process $\tm{P}$ is in canonical form if
  \[
    \tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}},
  \]
  such that: no process $\tm{P_i}$ is a cut; no process $\tm{P_i}$ is a link acting on a bound channel $\tm{x_i}$; and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel $\tm{x_i}$.
\end{definition}
\begin{corollary}
  If a process $\tm{P}$ is in canonical form, then it is blocked on an external communication.
\end{corollary}
\begin{proof}
  We have
  \[
  \tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}}
  \]
  such that no $\tm{P_i}$ is a cut or a link, and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel. The prefix of cuts introduces $n$ channels, and $n+1$ processes. Therefore, at least \emph{one} of the processes $\tm{P_i}$ must be acting on a free channel, i.e., blocked on an external communication.
\end{proof}
\begin{theorem}[Progress]\label{thm:cp-progress}
  If $\seq[P]{\Gamma}$, then either $\tm{P}$ is in canonical form, or there exists a process $\tm{Q}$ such that $\tm{P}\Longrightarrow\tm{Q}$.
\end{theorem} 
\begin{proof}
  We consider the maximum prefix of cuts of $\tm{P}$ such that $\tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}}$ and no $\tm{P_i}$ is a cut. If any process $\tm{P_i}$ is a link, we reduce by $(\cpLink{}{})$. If any two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same channel $\tm{x_i}$, we rewrite by $\equiv$ and reduce by the appropriate $\beta$-rule. Otherwise, $\tm{P}$ is in canonical form.
\end{proof}
\begin{theorem}[Termination]\label{thm:cp-termination}
  If $\seq[P]{\Gamma}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem} 
\begin{proof}
  Every reduction reduces a single cut to zero, one or two cuts. However, each of these cuts is smaller, measured in the size of the cut formula. Furthermore, each instance of the structural congruence preserves the size of the cut. Therefore, there cannot be an infinite $\Longrightarrow$-reduction sequence.
\end{proof}

@@

Typing judgements in \hcp associate processes with hyper-environments. \textsc{H-Mix} composes two processes in parallel, but remembers that they are independent in the sequent. \textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of at least two independent processes, and connects them, eliminating the vertical bar (the side condition $\tm{x} \not\in \ty{\mathcal{G}}$ ensures that sessions are binary). 
%This prevents self-locking processes, \eg $\tm{\piRecv{x}{}{\piBoundSend{x}{}{\piHalt}}}$.
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes that the process $\tm{P}$ consists of $n$ independent, but potentially entangled processes, each of which communicates according to its own protocol $\Gamma_i$. 
  Typing judgements can be constructed using the inference rules below. 
  \\[1\baselineskip]
  {Structural rules} %\vspace{-3ex}
  \begin{center}
    \hcpInfAx
    \hcpInfCut
  \end{center}
  \begin{center}
    \hcpInfMix
    \hcpInfHalt
  \end{center}
%  \vspace{.3ex}
  {Logical rules} %\vspace{-3ex}
  \begin{center}
    \hcpInfBoundTens
    \hcpInfParr
  \end{center}
  \begin{center}
    \hcpInfOne
    \hcpInfBot
  \end{center}
  \begin{center}
    \hcpInfPlus1
    \hcpInfPlus2
  \end{center}
  \begin{center}
    \hcpInfWith
  \end{center}
  \begin{center}
    \hcpInfNil
    \hcpInfTop
  \end{center}
\end{definition}
Note that the logical rules disallow hyperenvironments. If they did not, the following processes would be well-typed, even though they are stuck (unless we allow delayed actions as in \dhcp):
\begin{center}
  \begin{prooftree*}
    \noLine\AXC{$\vphantom{\Gamma}$}\UIC{$\vphantom{\Gamma}$}
    \NOM{H-Mix$_0$}
    \UIC{$\seq
      [\piHalt]
      {\emptyhypercontext}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{z}{}{\piHalt}]
      {\tmty{z}{\one}}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{x}{}{\piBoundSend{z}{}{\piHalt}}]
      {\tmty{x}{\one} \hsep \tmty{z}{\one}}$}
    \SYM{(\bot)}
    \UIC{$\seq
      [\piRecv{x}{}{\piBoundSend{x}{}{\piBoundSend{z}{}{\piHalt}}}]
      {\tmty{x}{\one} \hsep \tmty{x}{\bot}, \tmty{z}{\one}}$}
    \NOM{Cut}
    \UIC{$\seq
      [\piNew{x}{(\piRecv{x}{}{\piBoundSend{x}{}{\piBoundSend{z}{}{\piHalt}}})}]
      {\tmty{z}{\one}}$}
  \end{prooftree*}\hspace*{-1.75em}%
  \begin{prooftree*}
    \AXC{}
    \NOM{H-Mix$_0$}
    \UIC{$\seq
      [{\piHalt}]
      {\emptyhypercontext}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{y}{}{\piHalt}]
      {\tmty{y}{\one}}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}]
      {\tmty{x}{\one} \hsep \tmty{y}{\one}}$}
    \AXC{}
    \NOM{Ax}
    \UIC{$\seq
      [\cpLink{x}{z}]
      {\tmty{x}{\bot}, \tmty{z}{\one}}$}
    \SYM{(\bot)}
    \UIC{$\seq
      [\piRecv{y}{}{\cpLink{x}{z}}]
      {\tmty{x}{\bot}, \tmty{y}{\bot}, \tmty{z}{\one}}$}
    \NOM{H-Mix}
    \BIC{$\seq
      [(\piPar{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}{\piRecv{y}{}{\cpLink{x}{z}}})]
      {\tmty{x}{\one} \hsep \tmty{y}{\one} \hsep \tmty{x}{\bot}, \tmty{y}{\bot}, \tmty{z}{\one}}$}
    \NOM{Cut}
    \UIC{$\seq
      [\piNew{y}{(\piPar{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}{\piRecv{y}{}{\cpLink{x}{z}}})}]
      {\tmty{x}{\one} \hsep \tmty{x}{\bot}, \tmty{z}{\one}}$}
    \NOM{Cut}
    \UIC{$\seq
      [\piNew{x}{\piNew{y}{(\piPar{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}{\piRecv{y}{}{\cpLink{x}{z}}})}}]
      {\tmty{z}{\one}}$}
  \end{prooftree*}
\end{center}\vspace{.5\belowdisplayskip}

@@

\subsection{Metatheory}
\hcp enjoys subject reduction, termination, and progress.
\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal{G}}$.
\end{lemma} 
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal{G}}$.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}
\begin{definition}[Actions]
  A process $\tm{P}$ acts on $\tm{x}$ whenever $\tm{x}$ is free in the outermost
  term constructor of $\tm{P}$, \eg $\tm{\cpSend{x}{y}{P}{Q}}$ acts on $\tm{x}$
  but not on $\tm{y}$, and $\tm{\cpLink{x}{y}}$ acts on both $\tm{x}$ and $\tm{y}$.
  A process $\tm{P}$ is an action if it acts on some channel $\tm{x}$.
\end{definition}
\begin{definition}[Canonical forms]
  A process $\tm{P}$ is in canonical form if
 $
  \tm{P} \equiv \tm{\piNew{x_1}{\dots\piNew{x_n}{(P_1 \mid \dots \mid P_{n+m+1})}}},
  $
  such that: no process $\tm{P_i}$ is a cut or a mix; no process $\tm{P_i}$ is a link acting on a bound channel $\tm{x_i}$; and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel $\tm{x_i}$.
\end{definition}
Note that we have added the restriction ``acting on a bound channel'' to the case for links. This was not necessary for \cp, as all links in \cp act on at least one bound channel. Consequently, processes such as $\tm{\cpLink{x}{y}}$ and $\tm{(\piPar{\cpLink{x}{y}}{\cpLink{z}{w}})}$ are considered to be in canonical form. This is a generalisation of \cp, where $\tm{\cpLink{x}{y}}$ is considered to be in canonical form. If this is objectionable, the reduction system can be extended with identity expansion, expanding, \eg the process $\tm{\cpLink{x^{\ty{\bot}}}{y}}$ to $\tm{\cpWait{x}{\cpHalt{y}}}$.
\begin{corollary}
  If a process $\tm{P}$ is in canonical form, then it is blocked on an external communication.
\end{corollary}
\begin{proof}
  We have
  $
  \tm{P} \equiv \tm{\piNew{x_1}{\dots\piNew{x_n}{(P_1 \ppar \dots \ppar P_{n+m+1})}}},
  $
  such that no $\tm{P_i}$ is a cut or a link acting on a bound channel, and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel. The prefix of cuts and mixes introduces $n$ channels. Each application of cut requires an application of mix, so the prefix introduces $n+m+1$ processes. Therefore, at least $m+1$ of the processes $\tm{P_i}$ must be acting on a free channel, i.e., blocked on an external communication.
\end{proof}
\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\Gamma}$, then either $\tm{P}$ is in canonical form, or there exists a process $\tm{Q}$ such that $\tm{P}\Longrightarrow\tm{Q}$.
\end{theorem} 
\begin{proof}
  We consider the maximum prefix of cuts and mixes of $\tm{P}$ such that
  \[
  \tm{P} \equiv \tm{\piNew{x_1}{\dots\piNew{x_n}{(P_1 \ppar \dots \ppar P_{n+m+1})}}},
  \]
  and no $\tm{P_i}$ is a cut. If any process $\tm{P_i}$ is a link, we reduce by $(\cpLink{}{})$. If any two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same channel $\tm{x_i}$, we rewrite by $\equiv$ and reduce by the appropriate $\beta$-rule. Otherwise, $\tm{P}$ is in canonical form.
\end{proof}
\begin{theorem}[Termination]\label{thm:hcp-termination}
  If $\seq[P]{\mathcal{G}}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem} 
\begin{proof}
  As \cref{thm:cp-termination}.
\end{proof}

@@

@@

    \UIC{$\seq[\piNew{x}{(P \ppar Q)}]{\mathcal{G} \hsep \mathcal{H} \hsep \Gamma, \Delta}$}

@@

@@

      \UIC{$\seq[\piNew{x}{P}]{\mathcal{G} \hsep \Gamma, \Delta}$}

@@

Disentanglement is terminating, and confluent up to the associativity and commutativity of mixes.
\begin{lemma}[Disentangle]\label{lem:hcp-disentangle}
  If $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ in \hcp, then there exist processes $\tm{P_1}, \dots, \tm{P_n}$ in \cp such that $\seq[P_1]{\Gamma_1}, \dots, \seq[P_n]{\Gamma_n}$ and
  \[
    \begin{prooftree*}
      \AXC{\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}}
    \end{prooftree*}
    \moveMixDown^\star
    \begin{prooftree*}
      \AXC{$\seq[\mtf{P_1}]{\Gamma_1}$}
      \AXC{$\dots$}
      \AXC{$\seq[\mtf{P_n}]{\Gamma_n}$}
      \NOM{H-Mix$^\star$}
      \TIC{$\seq[(\mtf{P_1} \ppar \dots \ppar \mtf{P_n})]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$}
    \end{prooftree*}
  \]
\end{lemma} 
\begin{proof}
  We repeatedly apply the $\moveMixDown$-rules to the derivation $\rho$ to move the mixes downwards.
  There are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if a mix gets stuck above a cut, it forms a \cp cut;
  \item
    if a mix gets stuck above a $(\tens)$, it forms a \cp $(\tens)$;
  \item
    otherwise, it moves all the way to the bottom.
  \end{enumerate*}
  All applications of $(\one)$ are followed by an application of \textsc{H-Mix$_0$}, forming a \cp $(\one)$.
\end{proof}
An environment can be internalised as a type by collapsing it as a series of pars.
\begin{definition}\label{def:bigparr}
  \[
  \begin{array}{lcll}
    \ty{\bigparr(\emptycontext)}
    & = & \ty{\bot}
    \\
    \ty{\bigparr(\tmty{x_1}{A_1} , \dots , \tmty{x_n}{A_n})} 
     & = & \ty{A_1 \parr \dots \parr A_n}
     & \text{ if } n \geq 1
    \\
  \end{array}
  \]
\end{definition}\noindent
\begin{lemma}\label{lem:cp-bigparr}
  If $\seq{\Gamma}$ in \cp, then $\seq{\bigparr\Gamma}$ in \cp.
\end{lemma} 
\begin{proof}
  By repeated application of $(\parr)$.
\end{proof}\noindent
Furthermore, a hyper-environment can be internalised as a type by collapsing it as a series of tensors, where each constituent environment is internalised using $\ty{\bigparr}$. The empty hyper-environment $\ty{\emptyhypercontext}$ is internalised as the unit of tensor.
\begin{definition}\label{def:bigtens}
  \[
  \begin{array}{lcll}
    \ty{\bigtens(\emptyhypercontext)}
    & = & \ty{\one}
    \\
    \ty{\bigtens(\Gamma_1 \hsep \dots \hsep \Gamma_n)}
    & = & \ty{\bigparr\Gamma_1 \tens \dots \tens \bigparr\Gamma_n}
    & \text{ if } n \geq 1
  \end{array}
  \]
\end{definition}\noindent

>>>>

\subsection{Metatheory}
\cp enjoys subject reduction, termination, and progress~\cite{wadler2012,lindley2015}.
\begin{lemma}[Preservation for $\equiv$]\label{lem:cp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\Gamma}$ iff $\seq[Q]{\Gamma}$.
\end{lemma} 
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:cp-preservation}
  If $\seq[P]{\Gamma}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\Gamma}$.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}
\begin{definition}[Actions]
  A process $\tm{P}$ acts on $\tm{x}$ whenever $\tm{x}$ is free in the outermost
  term constructor of $\tm{P}$, \eg $\tm{\cpSend{x}{y}{P}{Q}}$ acts on $\tm{x}$
  but not on $\tm{y}$, and $\tm{\cpLink{x}{y}}$ acts on both $\tm{x}$ and $\tm{y}$.
  A process $\tm{P}$ is an action if it acts on some channel $\tm{x}$.
\end{definition}
\begin{definition}[Canonical forms]
  A process $\tm{P}$ is in canonical form if
  \[
    \tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}},
  \]
  such that: no process $\tm{P_i}$ is a cut; no process $\tm{P_i}$ is a link acting on a bound channel $\tm{x_i}$; and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel $\tm{x_i}$.
\end{definition}
\begin{corollary}
  If a process $\tm{P}$ is in canonical form, then it is blocked on an external communication.
\end{corollary}
\begin{proof}
  We have $\tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}}$
  such that no $\tm{P_i}$ is a cut or a link, and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel. The prefix of cuts introduces $n$ channels, and $n+1$ processes. Therefore, at least \emph{one} of the processes $\tm{P_i}$ must be acting on a free channel, i.e., blocked on an external communication.
\end{proof}
\begin{theorem}[Progress]\label{thm:cp-progress}
  If $\seq[P]{\Gamma}$, then either $\tm{P}$ is in canonical form, or there exists a process $\tm{Q}$ such that $\tm{P}\Longrightarrow\tm{Q}$.
\end{theorem} 
\begin{proof}
  We consider the maximum prefix of cuts of $\tm{P}$ such that $\tm{P} \equiv \tm{\cpCut{x_1}{P_1}{\dots\cpCut{x_n}{P_n}{P_{n+1}}\dots}}$ and no $\tm{P_i}$ is a cut. If any process $\tm{P_i}$ is a link, we reduce by $(\cpLink{}{})$. If any two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same channel $\tm{x_i}$, we rewrite by $\equiv$ and reduce by the appropriate $\beta$-rule. Otherwise, $\tm{P}$ is in canonical form.
\end{proof}
\begin{theorem}[Termination]\label{thm:cp-termination}
  If $\seq[P]{\Gamma}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem} 
\begin{proof}
  Every reduction reduces a single cut to zero, one or two cuts. However, each of these cuts is smaller, measured in the size of the cut formula. Furthermore, each instance of the structural congruence preserves the size of the cut. Therefore, there cannot be an infinite $\Longrightarrow$-reduction sequence.
\end{proof}

@@

Typing judgements in \hcp associate processes with hyper-environments. \textsc{H-Mix} composes two processes in parallel, but remembers that they are independent in the sequent. \textsc{H-Cut} and $(\tens)$ take as their premise a process which consists of at least two independent processes, and connects them, eliminating the vertical bar (the side condition $\tm{x} \not\in \ty{\mathcal{G}}$ ensures that sessions are binary). 
\begin{definition}[Typing judgements]\label{def:hcp}
  A typing judgement $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ denotes that the process $\tm{P}$ consists of $n$ independent, but potentially entangled processes, each of which communicates according to its own protocol $\Gamma_i$. 
  Typing judgements can be constructed using the inference rules below. 
  \\[1\baselineskip]
  {Structural rules} %\vspace{-3ex}
  \begin{center}
    \hcpInfAx
    \hcpInfCut
  \end{center}
  \begin{center}
    \hcpInfMix
    \hcpInfHalt
  \end{center}
%  \vspace{.3ex}
  {Logical rules} %\vspace{-3ex}
  \begin{center}
    \hcpInfBoundTens
    \hcpInfParr
  \end{center}
  \begin{center}
    \hcpInfOne
    \hcpInfBot
  \end{center}
  \begin{center}
    \hcpInfPlus1
    \hcpInfPlus2
  \end{center}
  \begin{center}
    \hcpInfWith
  \end{center}
  \begin{center}
    \hcpInfNil
    \hcpInfTop
  \end{center}
We shall omit the side condition $\tm{x} \not\in \ty{\mathcal{G}}$ in applications of rule \textsc{H-Cut} whenever $\ty{\mathcal{G}}$ is empty.
\end{definition}
Note that the logical rules disallow hyperenvironments. If they did not, the following processes would be well-typed, even though they are stuck (unless we allow delayed actions as in \dhcp):
\begin{center}
  \begin{prooftree*}
    \noLine\AXC{$\vphantom{\Gamma}$}\UIC{$\vphantom{\Gamma}$}
    \NOM{H-Mix$_0$}
    \UIC{$\seq
      [\piHalt]
      {\emptyhypercontext}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{z}{}{\piHalt}]
      {\tmty{z}{\one}}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{x}{}{\piBoundSend{z}{}{\piHalt}}]
      {\tmty{x}{\one} \hsep \tmty{z}{\one}}$}
    \SYM{(\bot)}
    \UIC{$\seq
      [\piRecv{x}{}{\piBoundSend{x}{}{\piBoundSend{z}{}{\piHalt}}}]
      {\tmty{x}{\one} \hsep \tmty{x}{\bot}, \tmty{z}{\one}}$}
    \NOM{H-Cut}
    \UIC{$\seq
      [\piNew{x}{(\piRecv{x}{}{\piBoundSend{x}{}{\piBoundSend{z}{}{\piHalt}}})}]
      {\tmty{z}{\one}}$}
  \end{prooftree*}\hspace*{-1.85em}%
  \begin{prooftree*}
    \AXC{}
    \NOM{H-Mix$_0$}
    \UIC{$\seq
      [{\piHalt}]
      {\emptyhypercontext}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{y}{}{\piHalt}]
      {\tmty{y}{\one}}$}
    \SYM{(\one)}
    \UIC{$\seq
      [\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}]
      {\tmty{x}{\one} \hsep \tmty{y}{\one}}$}
    \AXC{}
    \NOM{Ax}
    \UIC{$\seq
      [\cpLink{x}{z}]
      {\tmty{x}{\bot}, \tmty{z}{\one}}$}
    \SYM{(\bot)}
    \UIC{$\seq
      [\piRecv{y}{}{\cpLink{x}{z}}]
      {\tmty{x}{\bot}, \tmty{y}{\bot}, \tmty{z}{\one}}$}
    \NOM{H-Mix}
    \BIC{$\seq
      [(\piPar{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}{\piRecv{y}{}{\cpLink{x}{z}}})]
      {\tmty{x}{\one} \hsep \tmty{y}{\one} \hsep \tmty{x}{\bot}, \tmty{y}{\bot}, \tmty{z}{\one}}$}
    \NOM{H-Cut}
    \UIC{$\seq
      [\piNew{y}{(\piPar{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}{\piRecv{y}{}{\cpLink{x}{z}}})}]
      {\tmty{x}{\one} \hsep \tmty{x}{\bot}, \tmty{z}{\one}}$}
    \NOM{H-Cut}
    \UIC{$\seq
      [\piNew{x}{\piNew{y}{(\piPar{\piBoundSend{x}{}{\piBoundSend{y}{}{\piHalt}}}{\piRecv{y}{}{\cpLink{x}{z}}})}}]
      {\tmty{z}{\one}}$}
  \end{prooftree*}
\end{center}\vspace{.5\belowdisplayskip}

@@

\subsection{Metatheory}
\hcp enjoys subject reduction, termination, and progress.
\begin{lemma}[Preservation for $\equiv$]\label{lem:hcp-preservation-equiv}
  If $\tm{P}\equiv\tm{Q}$, then $\seq[P]{\mathcal{G}}$ iff $\seq[Q]{\mathcal{G}}$.
\end{lemma} 
\begin{proof}
  By induction on the derivation of $\tm{P}\equiv\tm{Q}$.
\end{proof}
\begin{theorem}[Preservation]\label{thm:hcp-preservation}
  If $\seq[P]{\mathcal{G}}$ and $\reducesto{P}{Q}$, then $\seq[Q]{\mathcal{G}}$.
\end{theorem} 
\begin{proof}
  By induction on the derivation of $\reducesto{P}{Q}$.
\end{proof}
\begin{definition}[Actions]
  A process $\tm{P}$ acts on $\tm{x}$ whenever $\tm{x}$ is free in the outermost
  term constructor of $\tm{P}$, \eg $\tm{\cpSend{x}{y}{P}{Q}}$ acts on $\tm{x}$
  but not on $\tm{y}$, and $\tm{\cpLink{x}{y}}$ acts on both $\tm{x}$ and $\tm{y}$.
  A process $\tm{P}$ is an action if it acts on some channel $\tm{x}$.
\end{definition}
\begin{definition}[Canonical forms]
  A process $\tm{P}$ is in canonical form if
 $
  \tm{P} \equiv \tm{\piNew{x_1}{\dots\piNew{x_n}{(P_1 \mid \dots \mid P_{n+m+1})}}},
  $
  such that: no process $\tm{P_i}$ is a cut or a mix; no process $\tm{P_i}$ is a link acting on a bound channel $\tm{x_i}$; and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel $\tm{x_i}$.
\end{definition}
Note that we have added the restriction ``acting on a bound channel'' to the case for links. This was not necessary for \cp, as all links in \cp act on at least one bound channel. Consequently, processes such as $\tm{\cpLink{x}{y}}$ and $\tm{(\piPar{\cpLink{x}{y}}{\cpLink{z}{w}})}$ are considered to be in canonical form. This is a generalisation of \cp, where $\tm{\cpLink{x}{y}}$ is considered to be in canonical form. If this is objectionable, the reduction system can be extended with identity expansion, expanding, \eg the process $\tm{\cpLink{x^{\ty{\bot}}}{y}}$ to $\tm{\cpWait{x}{\cpHalt{y}}}$.
\begin{corollary}
  If a process $\tm{P}$ is in canonical form, then it is blocked on an external communication.
\end{corollary}
\begin{proof}
  We have
  $
  \tm{P} \equiv \tm{\piNew{x_1}{\dots\piNew{x_n}{(P_1 \ppar \dots \ppar P_{n+m+1})}}},
  $
  such that no $\tm{P_i}$ is a cut or a link acting on a bound channel, and no two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same bound channel. The prefix of cuts and mixes introduces $n$ channels. Each application of cut requires an application of mix, so the prefix introduces $n+m+1$ processes. Therefore, at least $m+1$ of the processes $\tm{P_i}$ must be acting on a free channel, i.e., blocked on an external communication.
\end{proof}
\begin{theorem}[Progress]\label{thm:hcp-progress}
  If $\seq[P]{\Gamma}$, then either $\tm{P}$ is in canonical form, or there exists a process $\tm{Q}$ such that $\tm{P}\Longrightarrow\tm{Q}$.
\end{theorem} 
\begin{proof}
  We consider the maximum prefix of cuts and mixes of $\tm{P}$ such that
  $\tm{P} \equiv \tm{\piNew{x_1}{\dots\piNew{x_n}{(P_1 \ppar \dots \ppar P_{n+m+1})}}}$,
  and no $\tm{P_i}$ is a cut. If any process $\tm{P_i}$ is a link, we reduce by $(\cpLink{}{})$. If any two processes $\tm{P_i}$ and $\tm{P_j}$ are acting on the same channel $\tm{x_i}$, we rewrite by $\equiv$ and reduce by the appropriate $\beta$-rule. Otherwise, $\tm{P}$ is in canonical form.
\end{proof}
\begin{theorem}[Termination]\label{thm:hcp-termination}
  If $\seq[P]{\mathcal{G}}$, then there are no infinite $\Longrightarrow$-reduction sequences.
\end{theorem} 
\begin{proof}
  As \cref{thm:cp-termination}.
\end{proof}

@@

    \AXC{$\tm{x} \not\in \ty{(\mathcal{G} \hsep \mathcal{H})}$}

@@

    \BIC{$\seq[\piNew{x}{(P \ppar Q)}]{\mathcal{G} \hsep \mathcal{H} \hsep \Gamma, \Delta}$}

@@

      \AXC{$\tm{x} \not\in \ty{\mathcal{G}}$}

@@

      \BIC{$\seq[\piNew{x}{P}]{\mathcal{G} \hsep \Gamma, \Delta}$}

@@

Disentanglement is terminating, and confluent up to the associativity and commutativity of mixes.
\begin{lemma}[Disentangle]\label{lem:hcp-disentangle}
  If $\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$ in \hcp, then there exist processes $\tm{P_1}, \dots, \tm{P_n}$ in \cp such that $\seq[P_1]{\Gamma_1}, \dots, \seq[P_n]{\Gamma_n}$ and
  \[
    \begin{prooftree*}
      \AXC{\seq[P]{\Gamma_1 \hsep \dots \hsep \Gamma_n}}
    \end{prooftree*}
    \moveMixDown^\star
    \begin{prooftree*}
      \AXC{$\seq[\mtf{P_1}]{\Gamma_1}$}
      \AXC{$\dots$}
      \AXC{$\seq[\mtf{P_n}]{\Gamma_n}$}
      \doubleLine
      \NOM{H-Mix}
      \TIC{$\seq[(\mtf{P_1} \ppar \dots \ppar \mtf{P_n})]{\Gamma_1 \hsep \dots \hsep \Gamma_n}$}
    \end{prooftree*}
  \]
\end{lemma} 
\begin{proof}
  We repeatedly apply the $\moveMixDown$-rules to the derivation $\rho$ to move the mixes downwards.
  There are three cases:
  \begin{enumerate*}[label={\alph*)}]
  \item
    if a mix gets stuck above a cut, it forms a \cp cut;
  \item
    if a mix gets stuck above a $(\tens)$, it forms a \cp $(\tens)$;
  \item
    otherwise, it moves all the way to the bottom.
  \end{enumerate*}
  All applications of $(\one)$ are followed by an application of \textsc{H-Mix$_0$}, forming a \cp $(\one)$.
\end{proof}
An environment can be internalised as a type by collapsing it as a series of pars.
\begin{definition}\label{def:bigparr}
  \[
  \begin{array}{lcll}
    \ty{\bigparr(\emptycontext)}
    & = & \ty{\bot}
    \\
    \ty{\bigparr(\tmty{x_1}{A_1} , \dots , \tmty{x_n}{A_n})} 
     & = & \ty{A_1 \parr \dots \parr A_n}
     & \text{ if } n \geq 1
    \\
  \end{array}
  \]
\end{definition}\noindent
\begin{lemma}\label{lem:cp-bigparr}
  If $\seq{\Gamma}$ in \cp, then $\seq{\bigparr\Gamma}$ in \cp.
\end{lemma} 
\begin{proof}
  By repeated application of $(\parr)$.
\end{proof}\noindent
Furthermore, a hyper-environment can be internalised as a type by collapsing it as a series of tensors, where each constituent environment is internalised using $\ty{\bigparr}$. The empty hyper-environment $\ty{\emptyhypercontext}$ is internalised as the unit of tensor.
\begin{definition}\label{def:bigtens}
  \[
  \begin{array}{lcll}
    \ty{\bigtens(\emptyhypercontext)}
    & = & \ty{\one}
    \\
    \ty{\bigtens(\Gamma_1 \hsep \dots \hsep \Gamma_n)}
    & = & \ty{\bigparr\Gamma_1 \tens \dots \tens \bigparr\Gamma_n}
    & \text{ if } n \geq 1
  \end{array}
  \]
\end{definition}\noindent
